---
title: "Experiments_notebook"
author: "Aoji Tang"
date: "2024-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstudioapi)
library(stringr)
library(dplyr)
library(purrr)
library(ggplot2)
library(igraph)
library(reshape2)
```

## Data Loading and Processing

The folder structure of this project should be like:

```{plaintext}
├── Project
    ├── codes
    │   ├── starting_code.R
    │   └── trial_igraph.R
    ├── data
    │   ├── Introducing_the_Enron_Corpus.pdf
    │   ├── enron_mail_20150507
    │   │   └── maildir
    │   ├── enron_mail_20150507.tar.gz
    │   └── klimt-ecml04.pdf
    ├── figures
    │   ├── email_cluster.png
    │   ├── email_cluster_all.png
    │   ├── email_network.png
    │   └── email_network_all.png
    ├── final_project.Rmd
    ├── final_project.html
    ├── notebooks
    │   └── experiments.Rmd
    └── results
```

First, we set the paths of all the mail data. And the `mother_path` for
the project, containing data, code, notebooks, etc.

```{r}
# set the paths and working directory
notebook_path <- rstudioapi::getActiveDocumentContext()$path # get the file path for this notebook
mother_path <- dirname(dirname(notebook_path)) # since this notebook is under /notebooks, we just use its grandparent's path
mails_path <- paste0(mother_path, '/data/enron_mail_20150507/maildir/') # set the path for the folder containing all the mails
```

Retrieve all the **emails' filenames** and all of inbox emails' names in
the `maildir` of the enron dataset:

-   `emails.all.files`: store the **absolute** filepaths of all the
    emails in the enron dataset.

-   `emails.inbox.files`: store the absolute filepaths of all the
    **inbox** emails.

### inbox email data

```{r}
# Retrieval all files in the maildir

emails.all.files <- list.files(mails_path, full.names = T, recursive = T)
cat('The dataset contains', length(emails.all.files), 'emails in total')

# can save the files into the /results folder so that next time won't need to read the filenames again
# save(emails.all.files, file = paste0(mother_path, "/results/emails_all_files.Rdata")) 
# load(file = paste0(mother_path, "/results/emails_all_files.Rdata"))


# Filter by inbox emails only with regex

emails.inbox.files <- emails.all.files[grep("/inbox", emails.all.files)]
cat('The dataset contains', length(emails.inbox.files), 'inbox mails in total.')

# can save the files into the /results folder so that next time won't need to read the inbox email filenames again
# save(emails.inbox.files, file = paste0(mother_path, "/results/emails_inbox_files.Rdata"))
# load(file = paste0(mother_path, "/results/emails_inbox_files.Rdata"))
```

Filtering all the inbox emails within the company (by filtering all the
emails sent by another mail address ending with "\@enron.com". Then,
extracting all the **FROM** mailnames, and use the folder name as **TO**
entries.

```{r}
# Create list of sender and receiver (inbox owner)

inboxes.from.to.raw.df <- data.frame(
  from = apply(as.data.frame(emails.inbox.files), 1, function(x){readLines(x, warn = F)[3]}), # read all the FROM lines
  to = emails.inbox.files, # use the folder name for TO
  stringsAsFactors = F # remain the strings as characters
)

# Continue processing inboxes.from.to.raw.df

inboxes.from.to.df <- inboxes.from.to.raw.df %>%
  filter(grepl("@enron.com", from))  # Filter for inside-company emails by filtering all the address end with "@enron.com"

inboxes.from.to.df <- inboxes.from.to.df %>% mutate(
    from = str_sub(from, 7, nchar(from) - 10),  # Extract "FROM" email address by cutting "@enron.com" in the end
    to = sapply(str_split(inboxes.from.to.df$to, "/"), "[", 9)     # Extract "TO" name from folder name
  )

```

Create a dataframe of user folder names and filter by whether active or
not by detecting their "sent" folders. Then collect all the mailnames a
user would use (for example, (folder name:) allen-p has k..allen and
philips.allen both as his mailname). In the end, the `users` dataframe
contains three columns: `user.folder`, `mailname` and `name`(folder
name, such as allen-p).

**REMARK**: There is a sub-folder in the "/sent_items/" folder in
pereira-s named "clickathome", after checking the only email's content
in it (Ads), we choose to remove it from our investigation.

```{r}
# Create list of usernames in inboxes and remove those users without "sent mails" (inactive account)

users <- data.frame(user.folder = paste0(mails_path, unique(inboxes.from.to.df$to))) # add a user.folder column
sent <- apply(users, 1, function(x){sum(grepl("sent", dir(x)))}) # filter the users by active or not
users <- subset(users, sent != 0)

# Replace user.folder name with e-mail name

users <- users %>%
  mutate(mailname = vector("list", length = nrow(users)))

# There is a sub-folder in the "/sent_items/" folder in pereira-s named "clickathome", after checking the only email's content in it (ads), we choose to remove it from our investigation

all_possible_mailnames <- c() # store all the mailnames so that we can filter the mails by all the mailnames we have

for (i in 1:nrow(users)){
  
  # for each user, list all his folders
  all_folder_names <- list.dirs(users$user.folder[i], recursive=T, full.names=T)
  
  # get all folders named including "sent", so that we can have all the mailnames one uses
  sent_folders <- all_folder_names[grepl("sent", basename(all_folder_names), ignore.case = TRUE)]
  my_namelist <- c() # store the user's mailnames
  
  for (sent_folder in sent_folders){
    
    # get a list of all mails sent
    sentmail <- dir(sent_folder)
    
    # for each mail, extract the mailnames the user uses and store it as a list in the dataframe 'users'
    for (j in range(length(sentmail))){
      name <- readLines(paste0(sent_folder,'/', sentmail[j]), warn = F)[3] 
      name <- str_sub(name, 7, nchar(name)-10)
      
      # if the name is not in the list, then add it
      if (!(name %in% my_namelist)){ 
        my_namelist <- c(my_namelist, name)
        all_possible_mailnames <- c(all_possible_mailnames, name)
      }
      
    }
    
  }
  users$mailname[[i]] <- my_namelist # add the mailname list to one's row
}

# print(users$mailname)

# clean all the mailnames with regex, removing @ and special characters, leaving all the dots(.)
users <- users %>%
    mutate(
    mailname = map(mailname, ~ str_replace_all(.x, "[< >*@]", "")) %>%  # Apply regex to each element in the list
      map(~ str_replace_all(.x, "@$", ""))  # Apply second regex
  )

# get a column to connect the mailnames with the names.
users <- users %>%
  mutate(name = sapply(str_split(user.folder, '/'), '[', 8)) 

# users[1:10, ]
```

Further, we pick out all the inbox mails that are sent within all the
`users` we have til now. And we plot four histograms on the number of
mails received/sent with or without screening by at least 10.

```{r}
# filter the mails by asking both the sender and receivers are in our users list
inboxes.within.fromto.df <- inboxes.from.to.df %>%
  filter(from %in% all_possible_mailnames)
# print(inboxes.within.fromto.df)

# match the mailnames with names
inboxes.within.fromto.df <- inboxes.within.fromto.df %>%
  mutate(match_row = match(from, users$mailname)) %>%
  filter(!is.na(match_row)) %>%
  mutate(name = users$name[match_row], from = name) %>%
  select(-match_row, -name)
# print(inboxes.within.fromto.df)
```

Make exploratory plots:

```{r}
# make histograms of all the receivers and senders
hist_sent <- ggplot(inboxes.within.fromto.df, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of inbox emails sent within company", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hist_received <- ggplot(inboxes.within.fromto.df, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of inbox emails received within company", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)

# make plots for filtered data
filtered_sent <- inboxes.within.fromto.df %>%
  group_by(from) %>%
  filter(n() >= 20) %>%
  ungroup()

hist_sent <- ggplot(filtered_sent, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of inbox emails sent within company (>=20)", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

filtered_receive <- inboxes.within.fromto.df %>%
  group_by(to) %>%
  filter(n() >= 20) %>%
  ungroup()

hist_received <- ggplot(filtered_receive, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of inbox emails received within company(>=20)", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)
```

Also, we check the mail clusters and social network plots:

```{r}
set.seed(321)
# Get the unique senders and recipients
all_names <- unique(c(inboxes.within.fromto.df$from, inboxes.within.fromto.df$to))

# Create a contingency table with a predefined set of row and column names
mail_count_table <- table(factor(inboxes.within.fromto.df$from, levels = all_names), factor(inboxes.within.fromto.df$to, levels = all_names))

# print(table_inboxes_from_to_within)

mail_count_df <- melt(mail_count_table)

# Create a heatmap using ggplot2
ggplot(mail_count_df, aes(x = Var1, y = Var2, fill = log(value))) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Email Interaction Heatmap", x = "From", y = "To", fill = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create a graph from the matrix
graph <- graph_from_adjacency_matrix(mail_count_table, mode = "undirected", weighted = TRUE, diag = FALSE)
# Filter edges with low weight (e.g., below a threshold)
# graph <- delete_edges(graph, E(graph)[weight < 5])
community <- cluster_louvain(graph)
png(paste0(mother_path, "/figures/email_inbox_cluster.png"), width=3200, height = 3200)
plot(community, graph, vertex.size=4)
dev.off()

layout <- layout_with_fr(graph)  # Fruchterman-Reingold layout (often better for clarity)
png(paste0(mother_path, "/figures/email_inbox_network.png"), width = 3200, height = 2400)  # Width and height in pixels


# Plot the network graph with adjustments
plot(graph, 
     vertex.size = 5,         # Larger nodes
     vertex.label.cex = 0.8,  # Adjust text size
     edge.width = E(graph)$weight,  # Edge width based on the weight (email count)
     layout = layout,         # Use the new layout for better node spacing
     main = "Email Interaction Network",
     vertex.label.color = "black", # Change label color for contrast
     vertex.color = "lightblue",  # Node color
     edge.arrow.size = 0.5,   # Adjust arrow size on edges
     edge.color = "gray",     # Edge color
     vertex.label.dist = 1,   # Distance between label and node
     vertex.frame.color = "white")  # Frame color around nodes

# Close the PNG device (save the plot)
dev.off()
```

What happened to causholli-m? (he/she form a cluster him/herself.)

```{r}
causholli_inbox <- inboxes.from.to.raw.df%>%
  filter(to == 'causholli-m') %>%
  filter(from %in% all_possible_mailnames)
```

It is just one email he/she sent him/herself,and that's why she is the
only group.

------------------------------------------------------------------------

### all email data

**Now we focus on all the emails within the enron company**:

```{r}
# Create list of sender and receiver (inbox owner)

all.fromto.rawdf <- data.frame(
  from = apply(as.data.frame(emails.all.files), 1, function(x){readLines(x, warn = F)[3]}), # read all the FROM lines
  to = emails.all.files, # use the folder name for TO
  stringsAsFactors = F # remain the strings as characters
)

# Continue processing inboxes.from.to.raw.df

all.fromto.df <- all.fromto.rawdf %>%
  filter(grepl("@enron.com", from))  # Filter for inside-company emails by filtering all the address end with "@enron.com"

all.fromto.df <- all.fromto.df %>% mutate(
    from = str_sub(from, 7, nchar(from) - 10),  # Extract "FROM" email address by cutting "@enron.com" in the end
    to = sapply(str_split(all.fromto.df$to, "/"), "[", 9)     # Extract "TO" name from folder name
  )

# print(all.fromto.df)
```

Create a list of user folder names and filter by whether active or not
by detecting their "sent" folders. Then collect all the mailnames a user
would use (for example, allen-p has k..allen and philips.allen both as
his mailname). In the end, the `users` dataframe contains three columns:
`user.folder`, `mailname` and `name`(folder name, such as allen-p).

**REMARK**: There is a sub-folder in the "/sent_items/" folder in
pereira-s named "clickathome", after checking the only email's content
in it (ads), we choose to remove it from our investigation.

```{r}
# filter the mails by asking both the sender and receivers are in our users list
all.within.fromto.df <- all.fromto.df %>%
  filter(from %in% all_possible_mailnames)
# print(inboxes.within.fromto.df)

# match the mailnames with names
all.within.fromto.df <- all.within.fromto.df %>%
  mutate(match_row = match(from, users$mailname)) %>%
  filter(!is.na(match_row)) %>%
  mutate(name = users$name[match_row], from = name) %>%
  select(-match_row, -name)

# inboxes.within.fromto.df

# make histograms of all the receivers and senders
hist_sent <- ggplot(all.within.fromto.df, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of all emails sent within company", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hist_received <- ggplot(all.within.fromto.df, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of all emails received within company", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)

# make plots for filtered data
filtered_sent <- all.within.fromto.df %>%
  group_by(from) %>%
  filter(n() >= 500) %>%
  ungroup()

hist_sent <- ggplot(filtered_sent, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of all emails sent within company (>=500)", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

filtered_receive <- all.within.fromto.df %>%
  group_by(to) %>%
  filter(n() >= 500) %>%
  ungroup()

hist_received <- ggplot(filtered_receive, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of all emails received within company(>=500)", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)
```

```{r}
set.seed(123)
# Get the unique senders and recipients
all_names <- unique(c(all.within.fromto.df$from, all.within.fromto.df$to))

# Create a contingency table with a predefined set of row and column names
mail_count_table <- table(factor(all.within.fromto.df$from, levels = all_names), factor(all.within.fromto.df$to, levels = all_names))

# print(table_inboxes_from_to_within)

mail_count_df <- melt(mail_count_table)

# Create a heatmap using ggplot2
ggplot(mail_count_df, aes(x = Var1, y = Var2, fill = log(value))) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Email Interaction Heatmap", x = "From", y = "To", fill = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create a graph from the matrix
graph <- graph_from_adjacency_matrix(mail_count_table, mode = "undirected", weighted = TRUE, diag = FALSE)
# Filter edges with low weight (e.g., below a threshold)
# graph <- delete_edges(graph, E(graph)[weight < 5])
community <- cluster_louvain(graph)
png(paste0(mother_path, "/figures/email_all_cluster.png"), width=3200, height = 3200)
plot(community, graph, vertex.size=4)
dev.off()

layout <- layout_with_fr(graph)  # Fruchterman-Reingold layout (often better for clarity)
png(paste0(mother_path, "/figures/email_all_network.png"), width = 3200, height = 3200)  # Width and height in pixels


# Plot the network graph with adjustments
plot(graph, 
     vertex.size = 5,        # Larger nodes
     vertex.label.cex = 0.8,  # Adjust text size
     edge.width = E(graph)$weight,  # Edge width based on the weight (email count)
     layout = layout,         # Use the new layout for better node spacing
     main = "Email Interaction Network",
     vertex.label.color = "black", # Change label color for contrast
     vertex.color = "lightblue",  # Node color
     edge.arrow.size = 0.5,   # Adjust arrow size on edges
     edge.color = "gray",     # Edge color
     vertex.label.dist = 1,   # Distance between label and node
     vertex.frame.color = "white")  # Frame color around nodes

# Close the PNG device (save the plot)
dev.off()
```

### Look further into the clusters and packaging the data

From the email_all_cluster figure, we pick two best performing clusers (with human eye), get the members and export them into cluster1 and cluster2. Now under cluster1 and cluster2, we have the names, email file directories and mailnames of these users within.

```{R}
# first save the community results into .csv files
community_df <- data.frame(
  node = V(graph)$name,
  community = membership(community)
)
write.csv(community_df, paste0(mother_path, "/results/community_assignments.csv"), row.names = FALSE)

# get the membership information
membership(community)
target_community <- 10
target_nodes_1 <- V(graph)$name[membership(community) == target_community]
#print(target_nodes)  # Display the nodes in the target community

cluster1 <- users[users$name %in% target_nodes_1, ]

target_community <- 9
target_nodes_2 <- V(graph)$name[membership(community) == target_community]
# print(target_nodes)  # Display the nodes in the target community

cluster2 <- users[users$name %in% target_nodes_2, ]

c1.within.df <- all.within.fromto.df %>%
  filter(from %in% cluster1$name) %>%
  filter(to %in% cluster1$name)

c2.within.df <- all.within.fromto.df %>%
  filter(from %in% cluster2$name) %>%
  filter(to %in% cluster2$name)

c1.filtered.df <- c1.within.df %>%
  filter(from != to)
c1_count_table <- table(factor(c1.filtered.df$from, levels = target_nodes_1), factor(c1.filtered.df$to, levels=target_nodes_1))
c1_count_table

c2.filtered.df <- c2.within.df %>%
  filter(from != to)
c2_count_table <- table(factor(c2.filtered.df$from, levels = target_nodes_2), factor(c2.filtered.df$to, levels=target_nodes_2))
c2_count_table

graph <- graph_from_adjacency_matrix(c2_count_table, mode = "directed", weighted = TRUE, diag = FALSE)
layout <- layout_with_fr(graph)
plot(graph)
```
