---
title: "Experiments_notebook"
author: "Aoji Tang"
date: "2024-11-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rstudioapi)
library(stringr)
library(dplyr)
library(purrr)
library(ggplot2)
library(igraph)
library(reshape2)
```

First, we set the paths of all the mail data. And the `mother_path` for the project, containing data, code, notebooks, etc.

```{r}
# set the paths and working directory

path <- rstudioapi::getActiveDocumentContext()$path # get the file path for this notebook
mother_path <- dirname(dirname(path)) # since this notebook is under /notebooks, we just use its grandparent's path
mails_path <- paste0(mother_path, '/data/enron_mail_20150507/maildir/') # set the path for the folder containing all the mails
```

Retrieve all the email names and all of inbox emails' names.

```{r}
# Retrieval all files in the maildir

emails.all.files <- list.files(mails_path, full.names = T, recursive = T)
cat('The dataset contains', length(emails.all.files), 'emails in total')
save(emails.all.files, file = paste0(mother_path, "/results/emails_all_files.Rdata"))

# Filter by inbox only

emails.inbox.files <- emails.all.files[grep("/inbox", emails.all.files)]
cat('The dataset contains', length(emails.inbox.files), 'inbox mails in total.')
save(emails.inbox.files, file = paste0(mother_path, "/results/emails_inbox_files.Rdata"))
```

Filtering all the inbox emails within the company (by filtering all the emails sent by another mail address ending with "\@enron.com". Then, extracting all the **FROM** mailnames, and use the folder name as **TO** entries.

```{r}
# Create list of sender and receiver (inbox owner)

inboxes.from.to.raw.df <- data.frame(
  from = apply(as.data.frame(emails.inbox.files), 1, function(x){readLines(x, warn = F)[3]}), # read all the FROM lines
  to = emails.inbox.files, # use the folder name for TO
  stringsAsFactors = F # remain the strings as characters
)

# Continue processing inboxes.from.to.raw.df

inboxes.from.to.raw.df <- inboxes.from.to.raw.df %>%
  filter(grepl("@enron.com", from))  # Filter for inside-company emails by filtering all the address end with "@enron.com"

inboxes.from.to.raw.df <- inboxes.from.to.raw.df %>% mutate(
    from = str_sub(from, 7, nchar(from) - 10),  # Extract "FROM" email address by cutting "@enron.com" in the end
    to = sapply(str_split(inboxes.from.to.raw.df$to, "/"), "[", 9)     # Extract "TO" name from folder name
  )

```

Create a list of user folder names and filter by whether active or not by detecting their "sent" folders. Then collect all the mailnames a user would use (for example, allen-p has k..allen and philips.allen both as his mailname). In the end, the `users` dataframe contains three columns: `user.folder`, `mailname` and `name`(folder name, such as allen-p). 

**REMARK**: There is a sub-folder in the "/sent_items/" folder in pereira-s named "clickathome", after checking the only email's content in it (ads), we choose to remove it from our investigation.

```{r}
# Create list of usernames in inboxes and remove those users without "sent mails" (inactive account)

users <- data.frame(user.folder = paste0(mails_path, unique(inboxes.from.to.raw.df$to))) # add a user.folder column
sent <- apply(users, 1, function(x){sum(grepl("sent", dir(x)))}) # filter the users by active or not
users <- subset(users, sent != 0)

# Replace user.folder name with e-mail name

users <- users %>%
  mutate(mailname = vector("list", length = nrow(users)))

# There is a sub-folder in the "/sent_items/" folder in pereira-s named "clickathome", after checking the only email's content in it (ads), we choose to remove it from our investigation

all_possible_mailnames <- c() # store all the mailnames so that we can filter the mails by all the mailnames we have

for (i in 1:nrow(users)){
  
  # for each user, list all his folders
  all_folder_names <- list.dirs(users$user.folder[i], recursive=T, full.names=T)
  
  # get all folders named including "sent", so that we can have all the mailnames one uses
  sent_folders <- all_folder_names[grepl("sent", basename(all_folder_names), ignore.case = TRUE)]
  my_namelist <- c() # store the user's mailnames
  
  for (sent_folder in sent_folders){
    
    # get a list of all mails sent
    sentmail <- dir(sent_folder)
    
    # for each mail, extract the mailnames the user uses and store it as a list in the dataframe 'users'
    for (j in range(length(sentmail))){
      name <- readLines(paste0(sent_folder,'/', sentmail[j]), warn = F)[3] 
      name <- str_sub(name, 7, nchar(name)-10)
      
      # if the name is not in the list, then add it
      if (!(name %in% my_namelist)){ 
        my_namelist <- c(my_namelist, name)
        all_possible_mailnames <- c(all_possible_mailnames, name)
      }
      
    }
    
  }
  users$mailname[[i]] <- my_namelist # add the mailname list to one's row
}

# print(users$mailname)

# clean all the mailnames with regex, removing @ and special characters, leaving all the dots(.)
users <- users %>%
    mutate(
    mailname = map(mailname, ~ str_replace_all(.x, "[< >*@]", "")) %>%  # Apply regex to each element in the list
      map(~ str_replace_all(.x, "@$", ""))  # Apply second regex
  )

# get a column to connect the mailnames with the names.
users <- users %>%
  mutate(name = sapply(str_split(user.folder, '/'), '[', 8)) 

# users[1:10, ]
```

Further, we pick out all the inbox mails that are sent within all the `users` we have til now. And we plot four histograms on the number of mails received/sent with or without screening by at least 10.

```{r}
# filter the mails by asking both the sender and receivers are in our users list
inboxes.within.fromto.df <- inboxes.from.to.raw.df %>%
  filter(from %in% all_possible_mailnames)
# print(inboxes.within.fromto.df)

# match the mailnames with names
inboxes.within.fromto.df <- inboxes.within.fromto.df %>%
  mutate(match_row = match(from, users$mailname)) %>%
  filter(!is.na(match_row)) %>%
  mutate(name = users$name[match_row], from = name) %>%
  select(-match_row, -name)

# inboxes.within.fromto.df

# make histograms of all the receivers and senders
hist_sent <- ggplot(inboxes.within.fromto.df, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of inbox emails sent within company", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hist_received <- ggplot(inboxes.within.fromto.df, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of inbox emails received within company", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)

# make plots for filtered data
filtered_sent <- inboxes.within.fromto.df %>%
  group_by(from) %>%
  filter(n() >= 10) %>%
  ungroup()

hist_sent <- ggplot(filtered_sent, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of inbox emails sent within company (>=10)", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

filtered_receive <- inboxes.within.fromto.df %>%
  group_by(to) %>%
  filter(n() >= 10) %>%
  ungroup()

hist_received <- ggplot(filtered_receive, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of inbox emails received within company(>=10)", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)
```

```{r}
# Get the unique senders and recipients
all_names <- unique(c(inboxes.within.fromto.df$from, inboxes.within.fromto.df$to))

# Create a contingency table with a predefined set of row and column names
mail_count_table <- table(factor(inboxes.within.fromto.df$from, levels = all_names), factor(inboxes.within.fromto.df$to, levels = all_names))

# print(table_inboxes_from_to_within)

mail_count_df <- melt(mail_count_table)

# Create a heatmap using ggplot2
ggplot(mail_count_df, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  labs(title = "Email Interaction Heatmap", x = "From", y = "To", fill = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Create a graph from the matrix
graph <- graph_from_adjacency_matrix(mail_count_table, mode = "undirected", weighted = TRUE, diag = FALSE)
# Filter edges with low weight (e.g., below a threshold)
# graph <- delete_edges(graph, E(graph)[weight < 5])
community <- cluster_louvain(graph)
png("email_cluster.png", width=3200, height = 3200)
plot(community, graph)
dev.off()

layout <- layout_with_fr(graph)  # Fruchterman-Reingold layout (often better for clarity)
png("email_network.png", width = 3200, height = 2400)  # Width and height in pixels


# Plot the network graph with adjustments
plot(graph, 
     vertex.size = 5,        # Larger nodes
     vertex.label.cex = 0.8,  # Adjust text size
     edge.width = E(graph)$weight,  # Edge width based on the weight (email count)
     layout = layout,         # Use the new layout for better node spacing
     main = "Email Interaction Network",
     vertex.label.color = "black", # Change label color for contrast
     vertex.color = "lightblue",  # Node color
     edge.arrow.size = 0.5,   # Adjust arrow size on edges
     edge.color = "gray",     # Edge color
     vertex.label.dist = 1,   # Distance between label and node
     vertex.frame.color = "white")  # Frame color around nodes

# Close the PNG device (save the plot)
dev.off()
```

```{r}
library(pheatmap)
heatmap(table_inboxes_from_to_within)
```
What happened to causholli-m? (he/she form a cluster him/herself.)

```{r}
causholli_inbox <- inboxes.from.to.raw.df%>%
  filter(to == 'causholli-m') %>%
  filter(from %in% all_possible_mailnames)
```
It is just one email he/she sent him/herself,and that's why she is the only group.

----------------------------------------------------------------------------------

**Now we focus on all the emails within the enron company**:

```{r}
# Create list of sender and receiver (inbox owner)

all.fromto.rawdf <- data.frame(
  from = apply(as.data.frame(emails.all.files), 1, function(x){readLines(x, warn = F)[3]}), # read all the FROM lines
  to = emails.all.files, # use the folder name for TO
  stringsAsFactors = F # remain the strings as characters
)

# Continue processing inboxes.from.to.raw.df

all.fromto.rawdf <- all.fromto.rawdf %>%
  filter(grepl("@enron.com", from))  # Filter for inside-company emails by filtering all the address end with "@enron.com"

all.fromto.rawdf <- all.fromto.rawdf %>% mutate(
    from = str_sub(from, 7, nchar(from) - 10),  # Extract "FROM" email address by cutting "@enron.com" in the end
    to = sapply(str_split(all.fromto.rawdf$to, "/"), "[", 9)     # Extract "TO" name from folder name
  )

all.fromto.rawdf
```

Create a list of user folder names and filter by whether active or not by detecting their "sent" folders. Then collect all the mailnames a user would use (for example, allen-p has k..allen and philips.allen both as his mailname). In the end, the `users` dataframe contains three columns: `user.folder`, `mailname` and `name`(folder name, such as allen-p). 

**REMARK**: There is a sub-folder in the "/sent_items/" folder in pereira-s named "clickathome", after checking the only email's content in it (ads), we choose to remove it from our investigation.

```{r}
# filter the mails by asking both the sender and receivers are in our users list
all.within.fromto.df <- all.fromto.rawdf %>%
  filter(from %in% all_possible_mailnames)
# print(inboxes.within.fromto.df)

# match the mailnames with names
all.within.fromto.df <- all.within.fromto.df %>%
  mutate(match_row = match(from, users$mailname)) %>%
  filter(!is.na(match_row)) %>%
  mutate(name = users$name[match_row], from = name) %>%
  select(-match_row, -name)

# inboxes.within.fromto.df

# make histograms of all the receivers and senders
hist_sent <- ggplot(all.within.fromto.df, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of all emails sent within company", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

hist_received <- ggplot(all.within.fromto.df, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of all emails received within company", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)

# make plots for filtered data
filtered_sent <- all.within.fromto.df %>%
  group_by(from) %>%
  filter(n() >= 100) %>%
  ungroup()

hist_sent <- ggplot(filtered_sent, aes(x = from)) +
  geom_bar() +
  labs(title = "histogram of all emails sent within company (>=10)", x = 'Sent by:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

filtered_receive <- all.within.fromto.df %>%
  group_by(to) %>%
  filter(n() >= 100) %>%
  ungroup()

hist_received <- ggplot(filtered_receive, aes(x = to)) +
  geom_bar() +
  labs(title = "histogram of all emails received within company(>=10)", x = 'Sent to:') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot(hist_received)
plot(hist_sent)
```